---
title: "FL_data_analyzing"
author: "Paula Wu"
date: "12/7/2021"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

## Import the libraries ##

```{r, message=FALSE, include=FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(maps)
library(readxl)
library(plotly)
library(choroplethr)
library(choroplethrMaps)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r, message = FALSE}
drug_overdose = read_csv("./data/VSRR_Provisional_Drug_Overdose_Death_Counts.csv") %>% 
  janitor::clean_names()
state_level = c(state.name[1:8], "District of Columbia", state.name[9:32],"New York City", state.name[33:50])
drug_overdose_52 = 
  drug_overdose %>% 
  filter(!(state_name %in% c("United States"))) %>% 
  relocate(state_name) %>% 
  mutate(month = factor(month, levels = month.name), # change month and year to factor
         year = factor(year),
         state_name = factor(state_name, levels = state_level)) %>% 
  arrange(state_name) %>% 
  group_by(state_name, year) %>% 
  mutate(month = sort(month))

```

It's hard to find drug-specific data for those state missing specific drug types. Opioids data is much easier to find. We should probably focus on opioids.<br> 
We drew a map of the opioids to justify our choice.
```{r}
opioids_df = 
  drug_overdose_52 %>%  
  ungroup() %>% 
  select(state_name, year, month, indicator, data_value) %>% 
  filter(!(state_name %in% c("Alabama", "Arkansas", "Florida", "Idaho", "Louisiana", "Minnesota", "Nebraska", "North Dakota", "Pennsylvania")),
         str_detect(indicator, "T4")) %>% 
  mutate(opioids_yn = ifelse(str_detect(indicator, "opioids"), TRUE, FALSE)) %>% 
  group_by(year, month, opioids_yn) %>% 
  summarize(opioids_rate = mean(data_value, na.rm = TRUE))
opioids_df %>% 
  ungroup() %>% 
  ggplot(aes(x = month, y = opioids_rate, group = opioids_yn, color = opioids_yn))+
  geom_point()+
  geom_line()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 1))+
  facet_grid(.~year)
```
<br>If we decided to focus on opioids, this will make my life much easier.

## FL part:
Overall review:
```{r}
fl_death = 
  drug_overdose_52 %>% 
  filter(state_name %in% "Florida",
         indicator %in%  c("Number of Deaths", "Number of Drug Overdose Deaths")) %>% 
  select(year, month, indicator, deaths = data_value) %>% 
  pivot_wider(
    names_from = indicator,
    values_from = deaths
  ) %>% 
  janitor::clean_names() %>% 
  group_by(year, month) %>% 
  mutate(
    percent_overdose_death = number_of_drug_overdose_deaths / number_of_deaths
  )
 
```
```{r}
fl_death  %>% 
  ggplot(aes(x = month, y = percent_overdose_death, group = year, color = year))+
  geom_point()+
  geom_line()
```
```{r}
fl_death %>% 
  ggplot(aes(x = month, y = percent_overdose_death, group = NA))+
  geom_point()+
  geom_line()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 1))+
  facet_grid(.~year)
```
## Just opioids
We can also focus just on opioids 
Data source: cdc wonder, request by queries, Drug overdose deaths are identified using underlying cause-of-death codes from the Tenth Revision of ICD (ICD–10): X40–X44 (unintentional), X60–X64 (suicide), X85 (homicide), and Y10–Y14 (undetermined). Opioids identified by selecting T40.0-40.4, 40.6
```{r, message=FALSE, warning=FALSE}
# read in data
fl_opi_death = read_csv("./data/fl_opi_death_15_19.csv") %>% 
  janitor::clean_names() %>% 
  select(state_name = state, year, month, deaths)

fl_opi_death_add = read_csv("./data/fl_opi_death_20_21.csv") %>% 
  janitor::clean_names() %>% 
  select(state_name = occurrence_state, year = year_code, month, deaths) 

fl_opi_death = bind_rows(fl_opi_death, fl_opi_death_add) %>% 
  filter(!is.na(month)) %>% 
  separate(month, into = c("month", "useless"), sep = "\\,") %>% 
  select(-useless) %>% 
  mutate(month = substr(month, 1,3),
  month = month.name[match(str_to_title(month), month.abb)],
  month = factor(month, levels = month.name),
  year = factor(year)) %>% 
  select(-state_name)
```
data could be suppressed as indicated by the cdc website, the data value also supported by florida government

But anyway let's plot<br>
First by year, this is the death count
```{r}
fl_opi_death  %>% 
  ggplot(aes(x = month, y = deaths, group = year, color = year))+
  geom_point()+
  geom_line()
```
what about each month within a year
```{r}
fl_opi_death %>% 
  ggplot(aes(x = month, y = deaths, group = NA))+
  geom_point()+
  geom_line()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 1))+
  facet_grid(.~year)
```
It's climbing, every single month. The 2020 data is less probably because it's provisional and some data still under investigation

## Income 

Let's analyze the number of death vs median household income
```{r}
fl_eco_df = 
  read_csv("./data/median_household_income_fl.csv") %>% 
  janitor::clean_names() %>% 
  select(year, household_income_by_race, household_income_by_race_moe, geography) %>% 
  filter(
    str_detect(geography, "FL|United States|Florida"),
    year >= "2015") %>% 
  mutate(year = factor(year))


fl_eco_df %>% 
  mutate(text_label = str_c("Year: ", year, "\nMedian Household Income: $", household_income_by_race, 
                            "\nMargin of error: ± $", household_income_by_race_moe)) %>% 
  plot_ly(
    x = ~year, y = ~household_income_by_race, color = ~geography, text = ~text_label, 
    alpha = 0.5, type = "scatter", mode = "markers+lines", colors = "viridis", error_y = ~list(array = household_income_by_race_moe)) %>% 
  layout(
    title = "Median Household Income: FLorida vs. The U.S",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Median Household Income"))

```
Income and drug overdose death percent , by year
```{r}
#patchwork
income_drug_df = 
  fl_death %>% 
  ungroup() %>% 
  group_by(year) %>% 
  summarize(overdose_death_rate = sum(number_of_drug_overdose_deaths)/sum(number_of_deaths)) %>% 
  inner_join(., fl_eco_df %>% filter(geography %in% "Florida"))
income_drug_df %>% 
  ggplot(aes(x = household_income_by_race, y = overdose_death_rate, group = NA, color = year))+
  geom_point()+
  geom_line()
```


```{r, message=FALSE, warning=FALSE}
fl_county_df = 
  read_csv("./data/NCHS_-_Drug_Poisoning_Mortality_by_County__United_States.csv") %>% 
  janitor::clean_names() %>% 
  filter(state %in% "Florida") %>% 
  select(year, county, population, death_rate = model_based_death_rate) %>% 
  separate(county, into = c("county", "useless"), sep = " County") %>% 
  select(-useless) %>% 
  mutate(year = factor(year),
         county = str_to_lower(county)) %>% 
  relocate(county)
#fl_county_df
```
```{r, message=FALSE, warning=FALSE,echo=FALSE}
highlight_county = function(county_fips)
{
  data(county.map, package="choroplethrMaps", envir=environment())
  df = county.map[county.map$region %in% county_fips, ]
  geom_polygon(data=df, aes(long, lat, group = group), color = "yellow", fill = NA, size = 1)
}
add_text_county = function(county_fips){
  data(county.map, package="choroplethrMaps", envir=environment())
  df = county.map[county.map$region %in% county_fips, ]
  #geom_text(data=df, aes(mean(long), mean(lat), label = paste0(str_to_title(pull(county_fips, county)), " County\n", pull(county_fips, death_rate))), color = "white")
  geom_label(data=df, aes(mean(long), mean(lat), label = paste0(str_to_title(pull(county_fips, county)), " County\nDR: ", round(pull(county_fips, death_rate),2))), fill = "white", size = 2.5)
}
```
```{r, message=FALSE, warning=FALSE}
data(county.fips)
abc = county.fips %>% 
  separate(polyname, into = c("state", "county"), sep = "\\,") %>% 
  filter(state %in% "florida") %>% 
  select(-state) %>% 
  as_tibble()
```

### Counties change, 5-yr interval{.tabset}

#### 2003
```{r, message=FALSE, warning=FALSE}
year_select = 2003
start_county_df = left_join(fl_county_df,abc, by = "county") %>% 
  select(county, year, death_rate, fips) %>% 
  filter(year == year_select)

fips_add = c(12027, 12091, 12109, 12111)
a = 1
for (i in c(13,46,55,56)){
  start_county_df[i,4] = fips_add[a]
  a = a+1
}
start_county_df %>% 
  group_by(fips) %>% 
  mutate(fips = as.numeric(fips)) %>% 
  rename(region = fips,
         value = death_rate) %>% 
  county_choropleth(state_zoom = c("florida"),
                     legend = "death_rate")+
  highlight_county(start_county_df[which.max(pull(start_county_df, death_rate)),])+
  add_text_county(start_county_df[which.max(pull(start_county_df, death_rate)),])
```

#### 2008
```{r, message=FALSE, warning=FALSE}
year_select = 2008
start_county_df = left_join(fl_county_df,abc, by = "county") %>% 
  select(county, year, death_rate, fips) %>% 
  filter(year == year_select)

fips_add = c(12027, 12091, 12109, 12111)
a = 1
for (i in c(13,46,55,56)){
  start_county_df[i,4] = fips_add[a]
  a = a+1
}
start_county_df %>% 
  group_by(fips) %>% 
  mutate(fips = as.numeric(fips)) %>% 
  rename(region = fips,
         value = death_rate) %>% 
  county_choropleth(state_zoom = c("florida"),
                     legend = "death_rate")+
  highlight_county(start_county_df[which.max(pull(start_county_df, death_rate)),])+
  add_text_county(start_county_df[which.max(pull(start_county_df, death_rate)),])
```

#### 2013
```{r, message=FALSE, warning=FALSE}
year_select = 2013
start_county_df = left_join(fl_county_df,abc, by = "county") %>% 
  select(county, year, death_rate, fips) %>% 
  filter(year == year_select)

fips_add = c(12027, 12091, 12109, 12111)
a = 1
for (i in c(13,46,55,56)){
  start_county_df[i,4] = fips_add[a]
  a = a+1
}
start_county_df %>% 
  group_by(fips) %>% 
  mutate(fips = as.numeric(fips)) %>% 
  rename(region = fips,
         value = death_rate) %>% 
  county_choropleth(state_zoom = c("florida"),
                     legend = "death_rate")+
  highlight_county(start_county_df[which.max(pull(start_county_df, death_rate)),])+
  add_text_county(start_county_df[which.max(pull(start_county_df, death_rate)),])
```

#### 2018
```{r, message=FALSE, warning=FALSE, echo=FALSE}
year_select = 2018
start_county_df = left_join(fl_county_df,abc, by = "county") %>% 
  select(county, year, death_rate, fips) %>% 
  filter(year == year_select)

fips_add = c(12027, 12091, 12109, 12111)
a = 1
for (i in c(13,46,55,56)){
  start_county_df[i,4] = fips_add[a]
  a = a+1
}

start_county_df %>% 
  group_by(fips) %>% 
  mutate(fips = as.numeric(fips)) %>% 
  rename(region = fips,
         value = death_rate) %>% 
  county_choropleth(state_zoom = c("florida"),
                     legend = "death_rate")+
  highlight_county(start_county_df[which.max(pull(start_county_df, death_rate)),])+
  add_text_county(start_county_df[which.max(pull(start_county_df, death_rate)),])
```

Back to [homepage](index.html)