---
title: "NY_data_analyzing"
author: "Yuxuan Chen"
date: "12/7/2021"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

## Data cleaning

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(viridis)
library(dplyr)
library(plotly)
library(readr)
library(maps)
library(choroplethr)
library(choroplethrMaps)


knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	fig.width = 9, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
drug_overdose = read_csv("./data/VSRR_Provisional_Drug_Overdose_Death_Counts.csv") %>% 
  janitor::clean_names()

state_level = c(state.name[1:8], "District of Columbia", state.name[9:32],"New York City", state.name[33:50])

drug_overdose_52 = 
  drug_overdose %>% 
  filter(!(state_name %in% c("United States"))) %>% 
  relocate(state_name) %>% 
  mutate(month = factor(month, levels = month.name), # change month and year to factor
         year = factor(year),
         state_name = factor(state_name, levels = state_level)) %>% 
  arrange(state_name) %>% 
  group_by(state_name, year) %>% 
  mutate(month = sort(month)) 

nyc_df = 
  drug_overdose_52 %>%
  filter(state_name == "New York City")

drug_percent_specified = 
  nyc_df %>%
  ungroup() %>% 
  select(year, month, indicator, data_value) %>% 
  filter(indicator == "Percent with drugs specified")

# Missing data of death counts for specific drug types in year 2016, 2017
# Percent with drugs specified	for each month from 2015-2021 are high enough, all above 98%
```

## Number of Drug Overdose Deaths by Drug Categories in NYC

```{r}
nyc_df %>% 
  ungroup() %>% 
  filter(year %in% c("2016","2017")) %>% 
  select(year, month, indicator, data_value, predicted_value)


summarize_nyc_drug = 
  nyc_df %>% 
  ungroup() %>% 
  select(year, month, indicator, data_value) %>% 
  filter(!(indicator %in% c("Number of Deaths","Number of Drug Overdose Deaths","Percent with drugs specified"))) %>%
  filter(!(year %in% c("2016","2017"))) %>% 
  mutate(indicator = as.factor(indicator)) %>% 
  mutate(
    indicator = fct_reorder(indicator, data_value)
  ) 


summarize_nyc_drug %>% 
  summarize(n = sum(data_value)) %>% 
  arrange(desc(n))

summarize_nyc_drug %>% 
  ggplot(aes(x = indicator, y = data_value, fill = indicator)) +
  geom_violin(alpha = 0.5) +
  scale_x_discrete(labels = c("Psychostimulants \n with abuse potential \n (T43.6)", "Methadone \n (T40.3)", "Natural & \n semi-synthetic  \n opioids \n (T40.2)", "Natural & semi-\n synthetic opioids, \n incl. methadone \n (T40.2, T40.3)", "Heroin \n (T40.1)", "Cocaine \n (T40.5)", "Synthetic opioids, \n excl. \n methadone \n (T40.4)", "Natural, \n semi-synthetic, & \n synthetic opioids, \n incl. methadone \n (T40.2-T40.4)", "Opioids \n (T40.0-T40.4,\n T40.6)")) +
  guides(fill = guide_legend(nrow = 6, byrow = TRUE)) +
  labs(
    title = "Number of Drug Overdose Deaths by Drug Categories in NYC (2015 - 2021)",
    x = "Drug Categories",
    y = "Number of Drug Overdose Deaths",
    caption = "Data comes from VSRR_Provisional_Drug_Overdose_Death_Counts dataset."
  )
```

## Percent of Drug Overdose Deaths over Total Number of Deaths by Year in NYC: 

```{r}
nyc_drug_overdose_death_df =
  nyc_df %>% 
  ungroup() %>% 
  select(year, month, indicator, data_value) %>% 
  filter(indicator %in% c("Number of Deaths", "Number of Drug Overdose Deaths")) %>% 
  pivot_wider(
    names_from = indicator,
    values_from = data_value
  ) %>% 
  janitor::clean_names() %>% 
  group_by(year, month) %>% 
  mutate(
    percent_overdose_death = number_of_drug_overdose_deaths / number_of_deaths
  )

nyc_drug_overdose_death_df %>%
  ungroup() %>% 
  ggplot(aes(x = month, y = percent_overdose_death, color = year)) +
  geom_point() +
  geom_line(aes(group = year)) +
  labs(
    title = "Percent of Drug Overdose Deaths over Total Number of Deaths by Year in NYC",
    x = "Months",
    y = "Percent of Drug Overdose Deaths",
    caption = "Data comes from VSRR_Provisional_Drug_Overdose_Death_Counts dataset from 2015 to 2021)."
  )

```

```{r}
nyc_drug_overdose_death_df %>%
  ungroup() %>% 
  ggplot(aes(x = month, y = percent_overdose_death, group = NA)) +
  geom_point() +
  geom_line() +
  facet_grid(.~ year) +
  labs(
    title = "Percent of Drug Overdose Deaths over Total Number of Deaths by Year in NYC",
    x = "Months",
    y = "Percent of Drug Overdose Deaths",
    caption = "Data comes from VSRR_Provisional_Drug_Overdose_Death_Counts dataset from 2015 to 2021."
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```

## Number of Drug Overdose Deaths with Drug Categories by Year in NYC

```{r}
summarize_nyc_drug %>% 
  ggplot(aes(x = year, y = data_value, color = indicator)) +
  geom_point() +
  geom_line(aes(group = indicator)) +
  scale_color_viridis(discrete = TRUE, labels = c("Psychostimulants with abuse potential (T43.6)", "Methadone (T40.3)", "Natural & semi-synthetic opioids (T40.2)", "Natural & semi-synthetic opioids, incl. methadone (T40.2, T40.3)", "Heroin (T40.1)", "Cocaine (T40.5)", "Synthetic opioids, excl. methadone (T40.4)", "Natural, semi-synthetic, & synthetic opioids, incl. methadone (T40.2-T40.4)", "Opioids (T40.0-T40.4,T40.6)")) +
  guides(color = guide_legend(nrow = 6, byrow = TRUE)) +
  labs(
    title = "Number of Drug Overdose Deaths with Drug Categories by Year in NYC",
    x = "Drug Categories",
    y = "Number of Drug Overdose Deaths",
    caption = "Data comes from VSRR_Provisional_Drug_Overdose_Death_Counts dataset between 2015 and 2021."
  )
```

## Median Household Income: New York vs. The U.S

```{r}
ny_eco_df = 
  read_csv("./data/median_household_income_ny.csv") %>% 
  janitor::clean_names() %>% 
  select(year, household_income_by_race, household_income_by_race_moe, geography) %>% 
  filter(year >= "2015",
         !(geography  %in% c("New York-Newark-Jersey City, NY-NJ-PA", "New York"))) %>% 
  mutate( 
    geography = str_replace(geography, "New York, NY", "New York City"),
    geography = str_replace(geography, ", NY", ""),
    year = factor(year))

ny_eco_df %>% 
  mutate(text_label = str_c("Year: ", year, "\nMedian Household Income: $", household_income_by_race, 
                            "\nMargin of error: ± $", household_income_by_race_moe)) %>% 
  plot_ly(
    x = ~year, y = ~household_income_by_race, color = ~geography, text = ~text_label, 
    alpha = 0.5, type = "scatter", mode = "markers+lines", colors = "viridis",error_y = ~list(array = household_income_by_race_moe)) %>% 
  layout(
    title = "Median Household Income: New York vs. The U.S",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Median Household Income"))
```

## Income vs. Percent of Drug Overdose Death by Year in NYC

```{r}
income_drug_df_ny = 
  nyc_drug_overdose_death_df %>% 
    ungroup() %>% 
    group_by(year) %>% 
    summarize(overdose_death_rate = sum(number_of_drug_overdose_deaths)/sum(number_of_deaths)) %>% 
    inner_join(., ny_eco_df %>% filter(geography %in% "New York City"))

income_drug_df_ny %>% 
  ggplot(aes(x = household_income_by_race, y = overdose_death_rate, group = NA, color = year)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Income vs. Percent of Drug Overdose Death by Year in NYC",
    x = "Household Income by Race",
    y = "Percent of Drug Overdose Death"
  )
```

### Map of NYC Counties Drug Overdose Deathrate Change, 5-yr interval{.tabset}

```{r}
ny_county_df = 
  read_csv("./data/NCHS_-_Drug_Poisoning_Mortality_by_County__United_States.csv") %>% 
  janitor::clean_names() %>% 
  filter(state %in% "New York") %>% 
  select(year, county, population, model_based_death_rate) %>% 
  rename(death_rate = model_based_death_rate) %>% 
  mutate(
    county = str_replace(county, "County, NY", "")) %>% 
  mutate(year = factor(year),
         county = str_to_lower(county)) %>% 
  filter(str_detect(county, "bronx|queens|kings|new york|richmond")) %>% 
  mutate(county = str_replace(county, " $", "")) %>% 
  relocate(county)
ny_county_df
```

```{r}
data(county.fips)
nyc_fip = county.fips %>% 
  filter(str_detect(polyname, "new york")) %>% 
  mutate(
    polyname = str_replace(polyname, "new york,", "")) %>% 
  filter(str_detect(polyname, "bronx|queens|kings|new york|richmond")) %>% 
  rename(county = polyname) %>% 
  as.tibble()

ny_county_df = left_join(ny_county_df,nyc_fip, by = "county")
```

```{r}
highlight_county = function(county_fips)
{
  data(county.map, package="choroplethrMaps", envir=environment())
  df = county.map[county.map$region %in% county_fips, ]
  geom_polygon(data=df, aes(long, lat, group = group), color = "yellow", fill = NA, size = 1)
}
add_text_county = function(county_fips){
  data(county.map, package="choroplethrMaps", envir=environment())
  df = county.map[county.map$region %in% county_fips, ]
  #geom_text(data=df, aes(mean(long), mean(lat), label = paste0(str_to_title(pull(county_fips, county)), " County\n", pull(county_fips, death_rate))), color = "white")
  geom_label(data=df, aes(mean(long), mean(lat), label = paste0(str_to_title(pull(county_fips, county)), " County\nDR: ", round(pull(county_fips, death_rate),2))), fill = "white", size = 3)
}
```

#### 2003
```{r, message=FALSE, warning=FALSE}
year_select = 2003
start_county_df = ny_county_df %>% 
  select(county, year, death_rate, fips) %>% 
  filter(year == year_select)

start_deathrate_df = 
  start_county_df %>%
  rename(region = fips,
        value = death_rate) %>% 
  select(value, region)

county_choropleth(start_deathrate_df, title = "Drug Overdose Death Rates of Counties in New York City in 2003",
                    legend = "Death Rates",
                    county_zoom = start_deathrate_df$region) +
highlight_county(start_county_df[which.max(pull(start_county_df, death_rate)),]) +
add_text_county(start_county_df[which.max(pull(start_county_df, death_rate)),])
```

#### 2008
```{r, message=FALSE, warning=FALSE, echo=FALSE}
year_select = 2008
start_county_df = ny_county_df %>% 
  select(county, year, death_rate, fips) %>% 
  filter(year == year_select)

start_deathrate_df = 
  start_county_df %>%
  rename(region = fips,
        value = death_rate) %>% 
  select(value, region)

county_choropleth(start_deathrate_df, title = "Drug Overdose Death Rates of Counties in New York City in 2008",
                    legend = "Death Rates",
                    county_zoom = start_deathrate_df$region) +
highlight_county(start_county_df[which.max(pull(start_county_df, death_rate)),]) +
add_text_county(start_county_df[which.max(pull(start_county_df, death_rate)),])
```

#### 2013
```{r, message=FALSE, warning=FALSE, echo=FALSE}
year_select = 2013
start_county_df = ny_county_df %>% 
  select(county, year, death_rate, fips) %>% 
  filter(year == year_select)

start_deathrate_df = 
  start_county_df %>%
  rename(region = fips,
        value = death_rate) %>% 
  select(value, region)

county_choropleth(start_deathrate_df, title = "Drug Overdose Death Rates of Counties in New York City in 2013",
                    legend = "Death Rates",
                    county_zoom = start_deathrate_df$region) +
highlight_county(start_county_df[which.max(pull(start_county_df, death_rate)),]) +
add_text_county(start_county_df[which.max(pull(start_county_df, death_rate)),])
```

#### 2018
```{r, message=FALSE, warning=FALSE, echo=FALSE}
year_select = 2018
start_county_df = ny_county_df %>% 
  select(county, year, death_rate, fips) %>% 
  filter(year == year_select)

start_deathrate_df = 
  start_county_df %>%
  rename(region = fips,
        value = death_rate) %>% 
  select(value, region)

county_choropleth(start_deathrate_df, title = "Drug Overdose Death Rates of Counties in New York City in 2018",
                    legend = "Death Rates",
                    county_zoom = start_deathrate_df$region) +
highlight_county(start_county_df[which.max(pull(start_county_df, death_rate)),]) +
add_text_county(start_county_df[which.max(pull(start_county_df, death_rate)),])
```

Back to [homepage](index.html)